---
title: "CaseStudy1Deliverable"
author: "Rajat Chandna, Andy Ho, An Nguyen, Jodi Pafford, and Tori Wheelis"
date: "June 17, 2018"
output: 
  html_document:
    keep_md: yes
---
# Brewing Bros Brewery Report of the US Beer Market


## Introduction

The purpose of this study is to help "Brewing Bros"-- an international brewing company looking to expand to the United States-- assess market size and competition, and generally get a picture of the US brewing market, in order to determine their next brewery location and identify a featured beer. The data used in this study was obtained from "The National Beer and Wine Association" and includes 2410 US craft beers from 558 US breweries, with information on their name, location, beer type, acolohol content, and bitterness measurement (if available).

```{r, echo=TRUE}
#More information on the files, their cleanup, etc. can be found in the accompanying codebook.
```

## Data Background Information

Data for this project, obtained from The National Beer and Wine Association,which works with the largest beer distributors in the US, originated as two different datasets of the most popular and profitable beers (from data collected internally through their distribution centers as well as through distribution partners) in the US: one focused on specific beers and information about them, such as the brewery they came from, their name, and their alcohol content (Beers.csv) and a dataset of breweries who manufactured the beer, with information including where they are located (Breweries.csv). Both data sets had errors and duplications that were corrected or eliminated prior to beginning the research.

The dataset was parsed for duplicate beers and breweries, which were deleted, and misspellings of cities were corrected. After the two separate datasets were cleaned of duplicates and erroneous data lines, the two datasets were merged to create one large dataset, for the purposes of our report. This larger dataset can be found attached as an addendum to the report, labeled "master.file". 

##### Beers File Tidying
```{r beer, eco=TRUE}
#Tidying of Beers:
#read in Beers data set with correct character formatting.
Beers <- read.csv("Beers.csv", header=TRUE, fileEncoding = 'UTF-8', stringsAsFactors = F)
```

```{r}
#Remove duplicate beers if ALL columns but the Beer ID are the same. 
UniqueBeers <- Beers[!duplicated(Beers[c(1,3,4,5,6,7)]),]
```

```{r}
#Group beers by Brewery and Size
if(FALSE){
  byBrewery <- split(UniqueBeers, list(UniqueBeers$Ounces, UniqueBeers$Brewery_id))

# Match First 3 Chars
for(i in 1:length(byBrewery)){
    uniTempVec <- unique(byBrewery[[i]]$Name)
    tempListVec <- character()
    for(i in 1:length(uniTempVec)){
      subTobeSearched <- substr(uniTempVec[i], 1,3)
      subTobeSearched <- paste0("^",subTobeSearched)
      tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
      if(length(tempVec) > 1){
        tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
      }
    }

    unitempListVec <- unique(tempListVec)
    for(i in 1:length(unitempListVec)){
      temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
      if(length(temp2Vec) > 1){
        for(i in 1:length(temp2Vec)){
          cat(temp2Vec[i], file = "./First3matches.csv", append = T)
          if(i != length(temp2Vec)){
            cat(", ", file = "./First3matches.csv", append = T)
          }  
        }
        cat("\n", file = "./First3matches.csv", append = T)
      }
    }
  }

# Match Last 3 chars
  for(i in 1:length(byBrewery)){
    uniTempVec <- as.character(unique(byBrewery[[i]]$Name))
    tempListVec <- character()
    for(i in 1:length(uniTempVec)){
      subTobeSearched <- substr(uniTempVec[i], nchar(uniTempVec[i]) - 3, nchar(uniTempVec[i]))
      subTobeSearched <- paste0(subTobeSearched, "$")
      tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
      if(length(tempVec) > 1){
        tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
      }
    }

    unitempListVec <- unique(tempListVec)
    for(i in 1:length(unitempListVec)){
      temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
      if(length(temp2Vec) > 1){
        for(i in 1:length(temp2Vec)){
          cat(temp2Vec[i], file = "./Last3matches.csv", append = T)
          if(i != length(temp2Vec)){
            cat(", ", file = "./Last3matches.csv", append = T)
          }  
        }
        cat("\n", file = "./Last3matches.csv", append = T)
      }
    }
  }
}
```

```{r}
#Removed beers with identical ABU, IBU, Brewery_id, Ounces but slight differences in name
UniqueBeers <- subset(UniqueBeers, Name != "Ranger IPA (Current)")
UniqueBeers <- subset(UniqueBeers, Name != "Shift (1)")
UniqueBeers <- subset(UniqueBeers, Name != "Point Special (Current)")
UniqueBeers <- subset(UniqueBeers, Name != "Cherry Ale (1)")
UniqueBeers <- subset(UniqueBeers, Name != "Pleasure Town")
UniqueBeers <- subset(UniqueBeers, Name != "Farmer Ted's Cream Ale")
UniqueBeers <- subset(UniqueBeers, Name != "Dry Cider")
UniqueBeers <- subset(UniqueBeers, Name != "Abbey's Single (2015- )")
UniqueBeers <- subset(UniqueBeers, Name != "Triomphe")
UniqueBeers <- subset(UniqueBeers, Name != "Bender")
UniqueBeers <- subset(UniqueBeers, Name != "Hipster Ale (Westbrook Brewing)")

#rename Hipster Ale (Two Roads Brewing) to Hipster Ale; Two Roads Brewing does not brew Hipster Ale
levels(UniqueBeers$Name)[levels(UniqueBeers$Name) == "Hipster Ale (Two Roads Brewing)"] = "Hipster Ale"
```
```{r}
#Add code to change Brewery_ID column to match the Brews file and create CSV output of clean data.
colnames(UniqueBeers)[colnames(UniqueBeers)=="Brewery_id"] <- "Brew_ID"

#Write Clean File
write.csv(UniqueBeers, file="CleanedBeerData.csv", row.names = F)

```
##### Breweries File Tidying
```{r}
#Tidying Breweries.csv Section. Completed by Rajat and Andy
#First inspect the State column.  We listed out the States and got the count for each.  If observation > 50, we investigated further by examining all the States with a count of 1 (DC, ND, SD, WV).  In this case, there was 51 States and confirmed the 51st was DC which is acceptable.  All abbreviations in State column is valid.
breweries <- read.csv("Breweries.csv", header = TRUE)
str(breweries)
summary(breweries)
table(breweries$State)
```

```{r}
#In this section, looking to clean up the "City" column.  Still looking for a more elequent code but for now creating subsets for each letter and visually identifying misspellings, abbreviations, and/or punctuations.  Rajat to research a function that creates a subset of cities by subsetting cities with the same 3 letter string.
unique(breweries[grep("[.]", breweries$City), "City"])
unique(breweries[grep("^[Mm].*", breweries$City), "City"])
unique(breweries[grep("^[Ss].*", breweries$City), "City"])
```

```{r}
#Data Exploration Done, Cleaning begins   
BrDF <- read.csv("Breweries.csv", header = T, stringsAsFactors = F)
# First find and substitute any abbreviations symbols as "." with full form. Ex St. would be replaced
# with Saint, Mt. would be replaced with Mount etc
grep("[[:punct:]]",BrDF$City, value = T)
grep("[[:punct:][:blank:]]+",BrDF$City, value = T)
grep("^..[[:punct:][:blank:]]+",BrDF$City, value = T)
BrDF$City <- gsub("^St[[:punct:][:blank:]]+","Saint ",BrDF$City)
BrDF$City <- gsub("^Mt[[:punct:][:blank:]]+","Mount ",BrDF$City)

# The purpose of below code is to catch cities that could be misspelled or are located in 
# different state than what got entered by mistake.
# The logic is that if City Name starts with same characters, and later gets off by few charcters
# then the following code would catch such cities and list their states. If Thier States and Brewery Name
# Also Matches then these could be misspellings. The same loop would be repeated to match cities that end
# with same characters as these could be misspelled in their starting characters.

# Loop through every Unique city in City column, extract its first 3 characters,
# append ^ in front of it so as to make RegEx for the search. Search for all other
# Cities that start with Same 3 Characters. If more than 1 matching City is found
# append that city to running Vector(kind of appending to List)
uniTempVec <- unique(BrDF$City)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], 1,3)
  subTobeSearched <- paste0("^",subTobeSearched)
  tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

# List the States for such cities with matching first 3 chars. If state is also same,
# manually inspect brewery Names, to find potential misspellings.
unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(as.character(BrDF[BrDF$City == temp2Vec[j], "State"])))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) unique(BrDF[BrDF$City == x, "State"]))))
  }
}
```

```{r}
## Do corrections to City Names or State Names, If above code find issues.
# Corrections 
BrDF[BrDF$City == "Menominee", 'City'] <- "Menomonie"
BrDF[BrDF$City == "Menominie", 'City'] <- "Menomonie"
# Remove leading white space from State col
BrDF$State <- trimws(BrDF$State, "l")
# City Marquette is in MI and not MA
BrDF[BrDF$City == "Marquette" & BrDF$State == "MA", 'State'] <- "MI"
```

```{r}
# Now, repeat the above process to find potential missppelled cities based upon 
# Match of Last 3 Chars
uniTempVec <- unique(BrDF$City)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], nchar(uniTempVec[i]) - 3, nchar(uniTempVec[i]))
  subTobeSearched <- paste0(subTobeSearched,"$")
  tempVec <- unique(grep(subTobeSearched, uniTempVec,ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(as.character(BrDF[BrDF$City == temp2Vec[j], "State"])))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) unique(BrDF[BrDF$City == x, "State"]))))
  }
}

# No corrections after last 3 match

# Repeating above code to find cities with initial 2 chars match
uniTempVec <- unique(BrDF$City)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], 1,2)
  subTobeSearched <- paste0("^",subTobeSearched)
  tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(as.character(BrDF[BrDF$City == temp2Vec[j], "State"])))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) unique(BrDF[BrDF$City == x, "State"]))))
  }
}

# No significant findings when compared to 3 char Match, stopping with this City Column cleaning.
```

```{r}
# List any duplicate Rows based upon Name","City","State" columns.
BrDF[duplicated(BrDF[,c("Name","City","State")]), ]
# Remove Dups, if needed
#BrDF <- BrDF[!duplicated(BrDF[,c("Name","City","State")]), ]
```

```{r}
## Look for duplicates in Brewery Name
as.data.frame(table(BrDF$Name))[as.data.frame(table(BrDF$Name))$Freq > 1, ]
## Match first 3, same logic that was done to find misspelled cities.
uniTempVec <- unique(BrDF$Name)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], 1,3)
  subTobeSearched <- paste0("^",subTobeSearched)
  tempVec <- unique(grep(subTobeSearched, uniTempVec,ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- trimws(as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   "))), "l")
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(paste(BrDF[BrDF$Name == temp2Vec[j], "City"], BrDF[BrDF$Name == temp2Vec[j], "State"], sep = "+")))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) paste(BrDF[BrDF$Name == x, "City"], BrDF[BrDF$Name == x, "State"], sep = "+"))))
  }
}
```

```{r}
# Do corrections here
BrDF[BrDF$Name == "Against the Grain Brewery", "Name"] <- "Against The Grain Brewery"
BrDF[BrDF$Name == "Grey Sail Brewing of Rhode Island", "Name"] <- "Grey Sail Brewing Company"
BrDF[BrDF$Name == "Hops & Grains Brewing Company", "Name"] <- "Hops & Grain Brewery"
BrDF[BrDF$Name == "Goose Island Brewery Company", "Name"] <- "Goose Island Brewing Company"
BrDF[BrDF$Name == "Goodlife Brewing Co.", "Name"] <- "Good Life Brewing Company"
BrDF[BrDF$Name == "Catawba Brewing Company", "Name"] <- "Catawba Valley Brewing Company"
BrDF[BrDF$Name == "Angry Minnow", "Name"] <- "Angry Minnow Brewing Company"

#Write Clean File
write.csv(BrDF, file = "CleanedBreweryData.csv", row.names = F)
```

### What information do we have to work with?
The two separate files, beers and breweries, were merged into one large dataset. Some of the observations are seen below.  The data gathered includes... 

```{r Q2, echo=TRUE}
#Q2 - Merge Beer and Breweries files by Brew_ID (Assumes Beers.csv file has renamed Brewery_ID to Brew_ID) and prints the first/last 6 observations.
beers <- read.csv("CleanedBeerData.csv", header = TRUE)
breweries <- read.csv("CleanedBreweryData.csv", header = TRUE)
master.file <- Reduce(function(beers, breweries) merge(beers, breweries, by="Brew_ID"), list(beers, breweries))
names(master.file) <- c("Brewery_id", "Beer_name", "Beer_id", "ABV", "IBU", "Style", "Ounces", "Brewery_name", "City", "State")
write.csv(master.file, file = "masterfile.csv", row.names = FALSE)
head(master.file,6)
tail(master.file,6)
```

### Are there any obstacles with the data we're using?
Once the data was cleaned and combined, there were still some missing data (or data showing as "NA" or "incomplete".  Not all breweries keep all of these statistics on all of their artisan beers, especially IBU. These missing data include 997 IBU ratings and 62 ABV marks, as seen below. 

Only about 2.4% of the beers we have are missing information on alcohol by volume, although roughly 40% of the beers are missing information about their bitterness, using the International Bitterness Unit Scale. This still leaves us with a majority of our beers with complete information.

```{r Q3, echo = TRUE}
#Q3 - Summing the NA's for each column
colSums(is.na(master.file))

62/2410

997/2410
```

## Beer and Brewery Quick Facts

#### How many breweries are there per state? Are there any states that stand out as having too many or too few breweries?
The 558 breweries in this dataset are spread among all 50 states and includes District of Columbia (for a total of 51).  The most breweries are located in in Colorado (259), which has a booming beer market.  Other states with at least 100 breweries (and well-performing markets) include: Pennsylvania (100), Oregon (114), Texas (130), Indiana (138), Michigan (162), and California (181).

The states with the fewest breweries in the US also tend to be the smallest or least populated, such as Delaware (2), West Virginia (2), North Dakota (3), South Dakota (7), Washington, D.C. (8), Hew Hampshire (8), and New Jersey (8). However, a few others stand out, as they are larger or more populated, such as Arkansas (5), Tennessee (6), Alabama (10), Missouri (11), South Carolina (14), and Louisiana (19).

```{r Q1, echo=TRUE}
state.count <- as.data.frame(table(master.file$State))
names(state.count) <- c("State", "Brewery_Count")
statesSort <- state.count[order(state.count$Brewery_Count),]
statesSort
#Add Graphic for Presentation Purposes
library(ggplot2)
StateCount <- ggplot(statesSort, aes(reorder(State, -Brewery_Count), Brewery_Count, color=State)) + geom_bar(stat = "Identity", width = .85) + labs(x = "State") + theme(legend.position = "none") + coord_flip()
StateCount
```

### What is the most alcoholic (ABV) and most bitter(IBU) beer in the US?
Based on the information gathered, Colorado has the beer with the highest alcohol by volume (.128), followed closely by Kentucky (.125), and also Indiana (.120) and New York (.100). Delaware sits at the back of the pack with a max ABV of .055, followed by Arkansas (.061), Tennessee (.062), and New Hampshire (.065).

Oregon has the highest rating of International Bitterness Units (IBU) compared to the other states (138), also followed closely by Virginia (135), and also followed by Massachussetts (130) and Ohio (126). The state with the lowest IBU is Arkansas (39), followed by Delaware (52), Louisiana (60), and Tennessee (61), putting Tennessee in the bottom 5 for Alcohol by volume as well as International Bitterness Units. 

A full list can be seen below.

Note: A large portion (roughly 40%) of the beers are missing IBU information.

```{r Q5, echo = TRUE}
#Q4-7 done by Tori
MaxABV <- aggregate(ABV~State, data=master.file, max)
MaxABV <- MaxABV[order(MaxABV$ABV),]
MaxABV

MaxIBU <- aggregate(IBU~State, data=master.file, max)
MaxIBU <- MaxIBU[order(MaxIBU$IBU),]
MaxIBU

#Tori
```

### What does the US climate look like concerning alcohol by volume?
There were a somewhat wide range of beer ABV's in this dataset, from a .055 (Delaware) to a much higher .128 (Colorado). Other states with low maximum ABVs are Arkansas (.061), Tennessee (.062), New Hampshire (.065), and North Dakota (.067). Other states with high maximum ABVs are Kentucky (.125), Indiana (.120), and New York (.100), with the next thirteen states with their maximum alcohol by volume at .099.  The median ABV in this dataset is .09 and the third quartile is .099, which indicates a sizable portion of the beers on the market contain roughly 10% or so alcohol by volume. 

Of the 2410 beers in the dataset, 62 of them were missing ABV information, which is marginally small (roughly 2.5%). 

```{r Q6, echo = TRUE}
summary(master.file$ABV)

MaxABV <- aggregate(ABV~State, data=master.file, max)
MaxABV <- MaxABV[order(MaxABV$ABV),]
head(MaxABV)
tail(MaxABV)

summary(MaxABV)

```


### What does the median alcohol content & International Bitterness Units (IBU) for each state look like?
While knowing which states have the highest possible ABV's or IBU's, it can also be helpful to see where their "median", or middle-of-the-pack values for each are.

The median amount of alcohol content and IBUs in each state, and a summary per state, is listed below. 

```{r Q4a, echo=TRUE}
#Make a new dataframe with only median IBU and ABV values from each state
medianIBUABV<-aggregate.data.frame(master.file[, 4:5], list(master.file$State), median, na.rm=TRUE)
colnames(medianIBUABV)[1] <- "State"

medianIBUABV
```

The state with the highest median IBU is Maine, with 61. Following Maine in the top 5 is West Virginia (57.5). Next, tied at 55, are Florida and Georgia, followed by Oregon (54).

The states with the lowest median are Wisconsin and Kansas, tied at 20. Following Wisconsin and Kansas is Arizona (20.5), Wyoming (21), and Hawaii (23). You can see these in a list below.

```{r, Q4b, echo=TRUE}
#print the top and bottom five states for each category of IBU and ABV
medianIBU <- medianIBUABV[,c(1,3)]
medianIBU <- medianIBU[order(medianIBU$IBU),]

medianABV <- medianIBUABV[,c(1,2)]
medianABV <- medianABV[order(medianABV$ABV),]

#these are the states with the top 5 highest median IBU's
head(medianIBU, 5)
#these are the states with the bottom 5 highest median IBU's
tail(medianIBU, 6)

#these are the states with the top 5 highest median ABV's
head(medianABV, 5)
#these are the states with the bottom 5 highest median ABV's
tail(medianABV, 5)
```

Here is a visual snapshot of the differing median IBU and ABV values per state. These bar plots are ordered so that the largest values (of either ABV or IBU) are on the bottom, with the smallest on the top, to make viewing 50 state's worth of values simpler. Take note of the differing state's orders.

```{r Q4c, echo = TRUE}
#plot the following graphs side-by-side
par(mfrow=c(1,2))

#make a barplot for median IBU
medianIBUplot <- ggplot(medianIBUABV, aes(reorder(State, -IBU), IBU, color=State)) + geom_bar(stat = "Identity", width = .85) + labs(x = "State") + theme(legend.position = "none") + coord_flip()
medianIBUplot

#make a barplot for median ABV
medianABVplot <- ggplot(medianIBUABV, aes(reorder(State, -ABV), ABV, color=State)) + geom_bar(stat = "Identity", width = .85) + labs(x = "State") + theme(legend.position = "none") + coord_flip()
medianABVplot

```

### Is there a relationship between IBU and ABV in the US?
While we had somewhat limited data on the IBU of the beers in our dataset, since it is not a required measurement for breweries to track for the purposes of quality control, we were able to construct a scatterplot to help us see if there is a relationship between the IBU and the ABV of a beer. 

As you can see in the scatter plot below, it appears that there is a moderate correlation between the ABV and IBU of a beer between an ABV of .05-.10.
```{r Q7, echo = TRUE}

ggplot(master.file, aes(x=ABV, y=IBU))+
  geom_point(shape=19, alpha=(1/4)) +
  theme_bw() +
  theme() +
  ggtitle("Relationship Between Beer Bitterness and its Alcoholic Content") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
  theme(axis.title = element_text(face="bold"))
```


# Summary
Depending on Brewing Bros' desired approach, it is clear that the US beer market is booming and Brewing Bros will find a good home for their newest brewery. Based on the available data, it is the opinion of TAJAR, Inc. that Brewing Bros would make a sound investment setting up a brewery in a larger, more populous state with a small number of breweries that have a clear need for more artisan beers or more alcohol by volume beers, especially if those states are near states with very large beer markets.

Our preliminary research has identified Tennessee and Arkansas as two prime possible locations, with Missouri and South Carolina as other possible sites. Arkansas and Tennessee are known for low property values and taxes, and Arkansas in particular is known for its proximity to large beer markets, such as Texas and Colorado, and small but robust beer market. Arkansas has few breweries and a lot of potential. With few beers being produced that are both tracked by International Bitterness Units and that also score high on IBUs, in addition to a lower median alcohol by volume content, Arkansas demonstrates a clear market need for a standout IPA or Stout, both bitter beers, and another, well-respected brewery to put down roots.