---
title: "CaseStudy1Deliverable"
author: "Rajat Chandna, Andy Ho, An Nguyen, Jodi Pafford, and Tori Wheelis"
date: "June 17, 2018"
output: 
  html_document:
    keep_md: yes
---

#Introduction

The purpose of this study is to help "Brewing Bros" assess the market size and competition, to help them decide where the next "Brewing Bros" brewery should be built and what type of beer it should brew. The data used in this study was derived from "The National Beer and Wine Association" and includes 2410 US craft beers from 558 US breweries. (see Codebook for specific details)

#Tidy Data Process
Data for this project came from two different sources, a dataset containing beers and information about them, such as the brewery they came from, their name, and their alcohol content (Beers.csv) and a dataset of breweries, with information including where they are located (Breweries.csv).  Both data sets had extensive errors and duplications that were correct or eliminated prior to beginning the research.

The dataset was parsed for duplicate beers and breweries, which were deleted, and misspellings cities were corrected and any information that needed to be merged, was then merged. 

#####Beers File
```{r beer, eco=TRUE}
#Tidying of Beers:
#read in Beers data set with correct character formatting.
Beers <- read.csv("Beers.csv", header=TRUE, fileEncoding = 'UTF-8', stringsAsFactors = F)
```

```{r}
#Remove duplicate beers if ALL columns but the Beer ID are the same. 
UniqueBeers <- Beers[!duplicated(Beers[c(1,3,4,5,6,7)]),]
```

```{r}
#Group beers by Brewery and Size
if(FALSE){
  byBrewery <- split(UniqueBeers, list(UniqueBeers$Ounces, UniqueBeers$Brewery_id))

# Match First 3 Chars
for(i in 1:length(byBrewery)){
    uniTempVec <- unique(byBrewery[[i]]$Name)
    tempListVec <- character()
    for(i in 1:length(uniTempVec)){
      subTobeSearched <- substr(uniTempVec[i], 1,3)
      subTobeSearched <- paste0("^",subTobeSearched)
      tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
      if(length(tempVec) > 1){
        tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
      }
    }

    unitempListVec <- unique(tempListVec)
    for(i in 1:length(unitempListVec)){
      temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
      if(length(temp2Vec) > 1){
        for(i in 1:length(temp2Vec)){
          cat(temp2Vec[i], file = "./First3matches.csv", append = T)
          if(i != length(temp2Vec)){
            cat(", ", file = "./First3matches.csv", append = T)
          }  
        }
        cat("\n", file = "./First3matches.csv", append = T)
      }
    }
  }

# Match Last 3 chars
  for(i in 1:length(byBrewery)){
    uniTempVec <- as.character(unique(byBrewery[[i]]$Name))
    tempListVec <- character()
    for(i in 1:length(uniTempVec)){
      subTobeSearched <- substr(uniTempVec[i], nchar(uniTempVec[i]) - 3, nchar(uniTempVec[i]))
      subTobeSearched <- paste0(subTobeSearched, "$")
      tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
      if(length(tempVec) > 1){
        tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
      }
    }

    unitempListVec <- unique(tempListVec)
    for(i in 1:length(unitempListVec)){
      temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
      if(length(temp2Vec) > 1){
        for(i in 1:length(temp2Vec)){
          cat(temp2Vec[i], file = "./Last3matches.csv", append = T)
          if(i != length(temp2Vec)){
            cat(", ", file = "./Last3matches.csv", append = T)
          }  
        }
        cat("\n", file = "./Last3matches.csv", append = T)
      }
    }
  }
}
```

```{r}
#Removed beers with identical ABU, IBU, Brewery_id, Ounces but slight differences in name
UniqueBeers <- subset(UniqueBeers, Name != "Ranger IPA (Current)")
UniqueBeers <- subset(UniqueBeers, Name != "Shift (1)")
UniqueBeers <- subset(UniqueBeers, Name != "Point Special (Current)")
UniqueBeers <- subset(UniqueBeers, Name != "Cherry Ale (1)")
UniqueBeers <- subset(UniqueBeers, Name != "Pleasure Town")
UniqueBeers <- subset(UniqueBeers, Name != "Farmer Ted's Cream Ale")
UniqueBeers <- subset(UniqueBeers, Name != "Dry Cider")
UniqueBeers <- subset(UniqueBeers, Name != "Abbey's Single (2015- )")
UniqueBeers <- subset(UniqueBeers, Name != "Triomphe")
UniqueBeers <- subset(UniqueBeers, Name != "Bender")
UniqueBeers <- subset(UniqueBeers, Name != "Hipster Ale (Westbrook Brewing)")

#rename Hipster Ale (Two Roads Brewing) to Hipster Ale; Two Roads Brewing does not brew Hipster Ale
levels(UniqueBeers$Name)[levels(UniqueBeers$Name) == "Hipster Ale (Two Roads Brewing)"] = "Hipster Ale"
```
```{r}
#Add code to change Brewery_ID column to match the Brews file and create CSV output of clean data.
colnames(UniqueBeers)[colnames(UniqueBeers)=="Brewery_id"] <- "Brew_ID"

#Write Clean File
write.csv(UniqueBeers, file="CleanedBeerData.csv", row.names = F)

```
#####Brewery
```{r}
#Tidying Breweries.csv Section. Completed by Rajat and Andy
#First inspect the State column.  We listed out the States and got the count for each.  If observation > 50, we investigated further by examining all the States with a count of 1 (DC, ND, SD, WV).  In this case, there was 51 States and confirmed the 51st was DC which is acceptable.  All abbreviations in State column is valid.
breweries <- read.csv("Breweries.csv", header = TRUE)
str(breweries)
summary(breweries)
table(breweries$State)
```

```{r}
#In this section, looking to clean up the "City" column.  Still looking for a more elequent code but for now creating subsets for each letter and visually identifying misspellings, abbreviations, and/or punctuations.  Rajat to research a function that creates a subset of cities by subsetting cities with the same 3 letter string.
unique(breweries[grep("[.]", breweries$City), "City"])
unique(breweries[grep("^[Mm].*", breweries$City), "City"])
unique(breweries[grep("^[Ss].*", breweries$City), "City"])
```

```{r}
#Data Exploration Done, Cleaning begins   
BrDF <- read.csv("Breweries.csv", header = T, stringsAsFactors = F)
# First find and substitute any abbreviations symbols as "." with full form. Ex St. would be replaced
# with Saint, Mt. would be replaced with Mount etc
grep("[[:punct:]]",BrDF$City, value = T)
grep("[[:punct:][:blank:]]+",BrDF$City, value = T)
grep("^..[[:punct:][:blank:]]+",BrDF$City, value = T)
BrDF$City <- gsub("^St[[:punct:][:blank:]]+","Saint ",BrDF$City)
BrDF$City <- gsub("^Mt[[:punct:][:blank:]]+","Mount ",BrDF$City)

# The purpose of below code is to catch cities that could be misspelled or are located in 
# different state than what got entered by mistake.
# The logic is that if City Name starts with same characters, and later gets off by few charcters
# then the following code would catch such cities and list their states. If Thier States and Brewery Name
# Also Matches then these could be misspellings. The same loop would be repeated to match cities that end
# with same characters as these could be misspelled in their starting characters.

# Loop through every Unique city in City column, extract its first 3 characters,
# append ^ in front of it so as to make RegEx for the search. Search for all other
# Cities that start with Same 3 Characters. If more than 1 matching City is found
# append that city to running Vector(kind of appending to List)
uniTempVec <- unique(BrDF$City)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], 1,3)
  subTobeSearched <- paste0("^",subTobeSearched)
  tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

# List the States for such cities with matching first 3 chars. If state is also same,
# manually inspect brewery Names, to find potential misspellings.
unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(as.character(BrDF[BrDF$City == temp2Vec[j], "State"])))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) unique(BrDF[BrDF$City == x, "State"]))))
  }
}
```

```{r}
## Do corrections to City Names or State Names, If above code find issues.
# Corrections 
BrDF[BrDF$City == "Menominee", 'City'] <- "Menomonie"
BrDF[BrDF$City == "Menominie", 'City'] <- "Menomonie"
# Remove leading white space from State col
BrDF$State <- trimws(BrDF$State, "l")
# City Marquette is in MI and not MA
BrDF[BrDF$City == "Marquette" & BrDF$State == "MA", 'State'] <- "MI"
```

```{r}
# Now, repeat the above process to find potential missppelled cities based upon 
# Match of Last 3 Chars
uniTempVec <- unique(BrDF$City)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], nchar(uniTempVec[i]) - 3, nchar(uniTempVec[i]))
  subTobeSearched <- paste0(subTobeSearched,"$")
  tempVec <- unique(grep(subTobeSearched, uniTempVec,ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(as.character(BrDF[BrDF$City == temp2Vec[j], "State"])))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) unique(BrDF[BrDF$City == x, "State"]))))
  }
}

# No corrections after last 3 match

# Repeating above code to find cities with initial 2 chars match
uniTempVec <- unique(BrDF$City)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], 1,2)
  subTobeSearched <- paste0("^",subTobeSearched)
  tempVec <- unique(grep(subTobeSearched, uniTempVec, ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   ")))
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(as.character(BrDF[BrDF$City == temp2Vec[j], "State"])))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) unique(BrDF[BrDF$City == x, "State"]))))
  }
}

# No significant findings when compared to 3 char Match, stopping with this City Column cleaning.
```

```{r}
# List any duplicate Rows based upon Name","City","State" columns.
BrDF[duplicated(BrDF[,c("Name","City","State")]), ]
# Remove Dups, if needed
#BrDF <- BrDF[!duplicated(BrDF[,c("Name","City","State")]), ]
```

```{r}
## Look for duplicates in Brewery Name
as.data.frame(table(BrDF$Name))[as.data.frame(table(BrDF$Name))$Freq > 1, ]
## Match first 3, same logic that was done to find misspelled cities.
uniTempVec <- unique(BrDF$Name)
tempListVec <- character()
for(i in 1:length(uniTempVec)){
  subTobeSearched <- substr(uniTempVec[i], 1,3)
  subTobeSearched <- paste0("^",subTobeSearched)
  tempVec <- unique(grep(subTobeSearched, uniTempVec,ignore.case = T, value = T))
  if(length(tempVec) > 1){
    tempListVec <- c(tempListVec, paste(" ", tempVec, collapse = " "))
  }
}

unitempListVec <- unique(tempListVec)
temp3Vec <- character()
for(i in 1:length(unitempListVec)){
  temp3Vec <- character()
  temp2Vec <- trimws(as.vector(unlist(strsplit(trimws(unitempListVec[i], "l"), "   "))), "l")
  for (j in 1:length(temp2Vec)){
    temp3Vec <- c(temp3Vec,unique(paste(BrDF[BrDF$Name == temp2Vec[j], "City"], BrDF[BrDF$Name == temp2Vec[j], "State"], sep = "+")))
  }
  if(length(unique(temp3Vec)) != length(temp3Vec)){
    print(as.data.frame(sapply(temp2Vec, function(x) paste(BrDF[BrDF$Name == x, "City"], BrDF[BrDF$Name == x, "State"], sep = "+"))))
  }
}
```

```{r}
# Do corrections here
BrDF[BrDF$Name == "Against the Grain Brewery", "Name"] <- "Against The Grain Brewery"
BrDF[BrDF$Name == "Grey Sail Brewing of Rhode Island", "Name"] <- "Grey Sail Brewing Company"
BrDF[BrDF$Name == "Hops & Grains Brewing Company", "Name"] <- "Hops & Grain Brewery"
BrDF[BrDF$Name == "Goose Island Brewery Company", "Name"] <- "Goose Island Brewing Company"
BrDF[BrDF$Name == "Goodlife Brewing Co.", "Name"] <- "Good Life Brewing Company"
BrDF[BrDF$Name == "Catawba Brewing Company", "Name"] <- "Catawba Valley Brewing Company"
BrDF[BrDF$Name == "Angry Minnow", "Name"] <- "Angry Minnow Brewing Company"

#Write Clean File
write.csv(BrDF, file = "CleanedBreweryData.csv", row.names = F)
```

#Guiding Questions

###First 6, Last 6
The files were merged and some of the observations are seen below.  The data gathered includes... 

```{r Q2, echo=TRUE}
#Q2 - Merge Beer and Breweries files by Brew_ID (Assumes Beers.csv file has renamed Brewery_ID to Brew_ID) and prints the first/last 6 observations.
beers <- read.csv("CleanedBeerData.csv", header = TRUE)
breweries <- read.csv("CleanedBreweryData.csv", header = TRUE)
master.file <- Reduce(function(beers, breweries) merge(beers, breweries, by="Brew_ID"), list(beers, breweries))
names(master.file) <- c("Brewery_id", "Beer_name", "Beer_id", "ABV", "IBU", "Style", "Ounces", "Brewery_name", "City", "State")
write.csv(master.file, file = "masterfile.csv", row.names = FALSE)
head(master.file,6)
tail(master.file,6)
```

###Number of NA's
Once the data was cleaned and combined, there were still some missing data (or data showing as "NA" or "incomplete".  Not all breweries keep all of these statistics on all of their artisan beers, especially IBU. These missing data include 997 IBU ratings and 62 ABV marks, as seen below.

```{r Q3, echo = TRUE}
#Q3 - Summing the NA's for each column
colSums(is.na(master.file))
```

##Beer and Brewery Quick Facts

###Breweries per state
The 558 breweries are spread among all 50 states and includes District of Columbia (for a total of 51).  The most breweries are located in in Colorado (259).  Other states with at least 100 breweries include: Pennsylvania (100), Oregon (114), Texas (130), Indiana (138), Michigan (162), and California (181).  
```{r Q1, echo=TRUE}
state.count <- as.data.frame(table(master.file$State))
names(state.count) <- c("State", "Brewery_Count")
statesSort <- state.count[order(state.count$Brewery_Count),]
statesSort
#Add Graphic for Presentation Purposes
library(ggplot2)
StateCount <- ggplot(statesSort, aes(reorder(State, -Brewery_Count), Brewery_Count, color=State)) + geom_bar(stat = "Identity", width = .85) + labs(x = "State") + theme(legend.position = "none") + coord_flip()
StateCount
```

###Maximum Alcoholic (ABV) and Most Bitter(IBU) Beer
Based on the information gathered, Colorado has the highest alcohol by volumne (.128), followed closely by Kentucky (.125), and also Indiana (.120) and New York (.100). Delaware sits at the back of the pack with a max ABV of .055, followed by Arkansas (.061), Tennessee (.062), and New Hampshire (.065).

Oregon has the highest rating of International Bitterness Units (IBU) compared to the other states (138), also followed closely by Virginia (135), and also followed by Massachussetts (130) and Ohio (126). The state with the lowest IBU is Arkansas (39), followed by Delaware (52), Louisiana (60), and Tennessee (61), putting Tennessee in the bottom 5 for Alcohol by volume as well as International Bitterness Units. 

A full list can be seen below.

```{r Q5, echo = TRUE}
#Q4-7 done by Tori
MaxABV <- aggregate(ABV~State, data=master.file, max)
MaxABV <- MaxABV[order(MaxABV$ABV),]
MaxABV

MaxIBU <- aggregate(IBU~State, data=master.file, max)
MaxIBU <- MaxIBU[order(MaxIBU$IBU),]
MaxIBU

#Tori
```

###ABV Summary
There were a wide range of beer ABV's in this dataset, from a (safe for pregnancy) .001 to a much higher .128, with the median at .057. Of the 2410 beers in the dataset, 62 of them were missing ABV information, which is marginally small. 

```{r Q6, echo = TRUE}
summary(master.file$ABV)

MaxABV <- aggregate(ABV~State, data=master.file, max)
MaxABV <- MaxABV[order(MaxABV$ABV),]
head(MaxABV)
tail(MaxABV)

MaxIBU <- aggregate(IBU~State, data=master.file, max)
MaxIBU <- MaxIBU[order(MaxIBU$IBU),]
head(MaxIBU)
tail(MaxIBU)

#Tori
```


###Median Alcohol Content & International Bitterness Units (IBU) for each state
While knowing which states have the highest possible ABV's or IBU's, it can also be helpful to see where their "median", or middle-of-the-pack values for each are.

The median amount of alcohol content and IBUs in each state is listed below. 

```{r Q4a, echo=TRUE}
#Make a new dataframe with only median IBU and ABV values from each state
medianIBUABV<-aggregate.data.frame(master.file[, 4:5], list(master.file$State), median, na.rm=TRUE)
colnames(medianIBUABV)[1] <- "State"

medianIBUABV
```

The state with the highest median IBU is Maine, with 61. Following Maine in the top 5 is West Virginia (57.5). Next, tied at 55, are Florida and Georgia, followed by Oregon (54).

The state with the lowest median is Hawaii, with a score of 23. Following Hawaii is Wyoming, with a score of 21, Airzona with 20.5, and Wisconsin and Kansas tied with 20.

```{r, Q4b, echo=TRUE}
#print the top and bottom five states for each category of IBU and ABV
medianIBU <- medianIBUABV[,c(1,3)]
medianIBU <- medianIBU[order(medianIBU$IBU),]

medianABV <- medianIBUABV[,c(1,2)]
medianABV <- medianABV[order(medianABV$ABV),]

#these are the states with the top 5 highest median IBU's
head(medianIBU, 5)
#these are the states with the bottom 5 highest median IBU's
tail(medianIBU, 6)

#these are the states with the top 5 highest median ABV's
head(medianABV, 5)
#these are the states with the bottom 5 highest median ABV's
tail(medianABV, 5)
```

Here is a visual snapshot of the differing median IBU and ABV values per state.

```{r Q4c, echo = TRUE}
#plot the following graphs side-by-side
par(mfrow=c(1,2))

#make a barplot for median IBU
medianIBUplot <- ggplot(medianIBUABV, aes(reorder(State, -IBU), IBU, color=State)) + geom_bar(stat = "Identity", width = .85) + labs(x = "State") + theme(legend.position = "none") + coord_flip()
medianIBUplot

#make a barplot for median ABV
medianABVplot <- ggplot(medianIBUABV, aes(reorder(State, -ABV), ABV, color=State)) + geom_bar(stat = "Identity", width = .85) + labs(x = "State") + theme(legend.position = "none") + coord_flip()
medianABVplot

# Tori
```

###IBU and ABV relationship
While we had somewhat limited data on the IBU of the beers in our dataset, since it is not a required measurement for breweries to track for the purposes of quality control, we were able to construct a scatterplot to help us see if there is a relationship between the IBU and the ABV of a beer. 

As you can see in the scatter plot below, it appears that there is a moderate correlation between the ABV and IBU of a beer between ABV of .05-.10.
```{r Q7, echo = TRUE}

ggplot(master.file, aes(x=ABV, y=IBU))+
  geom_point(shape=19, alpha=(1/4)) +
  theme_bw() +
  theme() +
  ggtitle("Relationship Between Beer Bitterness and its Alcoholic Content") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"))
  theme(axis.title = element_text(face="bold"))
```


#Summary